---
title: "12. OOP Base Types"
output: html_notebook
---

```{r include=FALSE}
source("set_up.R", local = TRUE)
```

- **S3** - simple and common - is used throughout base R
- **R6** - encapsulated - is used to escape Rs copy-on-modify semantics. Used in cases such as modelling data that exist independently of R such as web APIs
- **S4** - similar to S3 but more formal and strict - is a rigorous system that forces developers to think about the design. This is suitable for large projects with many contributor. S4 was a rewrite of S3 to make it more formal.

Process of finding the correct method for a given class is **method dispatch**.

In **encapsulated OOP**, methods belong to objects or classes, and method calls typically look like `object.method(arg1, arg2)`. This is called encapsulated because the object encapsulates both data (with fields) and behaviour (with methods), and is the paradigm found in most popular languages.

In **functional OOP**, methods belong to generic functions, and method calls look like ordinary function calls: `generic(object, arg2, arg3)`. This is called functional because from the outside it looks like a regular function call, and internally the components are also functions.

```{r eval=FALSE}
sloop::otype(1:10)
#> "base"

sloop::otype(mtcars)
#> "S3"

mle_obj <- stats4::mle(function(x = 1) (x - 2) ^ 2)
sloop::otype(mle_obj)
#> "S4"
```

## Base types

There is a difference between base object and OO objects in R. Base objects do not have a class, but OO objects do.

```{r eval=FALSE}
# A base object:
is.object(1:10)
#> [1] FALSE
sloop::otype(1:10)
#> [1] "base"

# An OO object
is.object(mtcars)
#> [1] TRUE
sloop::otype(mtcars)
#> [1] "S3"
```

```{r eval=FALSE}
attr(1:10, "class")
#> NULL

attr(mtcars, "class")
#> "data.frame"
```

Only OO objects have a class attribute, every object has a base type

```{r eval=FALSE}
typeof(1:10)
#> "integer"

typeof(mtcars)
#> "list"
```

```{r eval=FALSE}
typeof(NULL)
#> [1] "NULL"
typeof(1L)
#> [1] "integer"
typeof(1i)
#> [1] "complex"
```

```{r eval=FALSE}
typeof(mean)
#> [1] "closure"
typeof(`[`)
#> [1] "special"
typeof(sum)    
#> [1] "builtin"
```

```{r eval=FALSE}
typeof(globalenv())
#> [1] "environment"
```

```{r eval=FALSE}
typeof(quote(a))
#> [1] "symbol"
typeof(quote(a + 1))
#> [1] "language"
typeof(formals(mean))
#> [1] "pairlist"
```

Numeric can mean double or integer

```{r, eval=FALSE}
sloop::s3_class(1)
#> [1] "double"  "numeric"
sloop::s3_class(1L)
#> [1] "integer" "numeric"
```

```{r eval=FALSE}
typeof(factor("x"))
#> [1] "integer"
is.numeric(factor("x"))
#> [1] FALSE
```

